\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\graphicspath{{./}{sections/}{images/}}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{url}

\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{The Devils Behind OpenClaw: A Survey of Safety, Security, and Governance in Computer-Use Agents}

\author{
\IEEEauthorblockN{
Zejian Chen,
Zhanyuan Liu,
Xi Zhang$^{*}$\thanks{*Corresponding author.},
Zhaozhuo Li,
Litian~Zhang,
and Mengxiang Han
}
\IEEEauthorblockA{
Beijing University of Posts and Telecommunications, Beijing, China \\
Email: \{chenzejian, doubao, zhangx, lichaozhuo, zhanglitian\}@bupt.edu.cn, hanmx12@chinaunicom.cn
}
}

\maketitle

\begin{abstract}
Computer-use agents (CUAs), which perceive on-screen content, reason over user goals, and execute environment-grounded actions in desktop, web, and mobile settings, are advancing rapidly. At the same time, emerging open deployments such as OpenClaw have made these capabilities more visible outside benchmark settings, drawing broader attention to the safety, security, and governance risks of computer-use agents in real-world environments. However, existing research on CUAs remains fragmented across architectural design, evaluation, deployment practice, and risk analysis. In this survey, we provide a unified view of the field through a three-layer architecture---Perception, Decision, and Execution---that captures how CUAs represent state, reason about tasks, and ground actions in interactive environments. We complement this architectural perspective with a lifecycle view spanning creation, deployment, execution, and maintenance, showing how capabilities, risks, and responsibilities evolve from model development to real-world operation. Building on these foundations, we organize the emerging threat landscape of CUAs, including prompt injection, interface deception, tool and privilege misuse, memory compromise, and broader supply-chain failures, and relate these risks to concrete control points across the system lifecycle. We further discuss governance mechanisms for CUA ecosystems, including evaluation and benchmarking, auditing and provenance, permission control and sandboxing, and cross-stakeholder accountability. Overall, this survey aims to clarify how computer-use agents are built, where they fail, and what is required to make them more trustworthy, controllable, and sustainable at scale.
\end{abstract}

\begin{IEEEkeywords}
computer-use agents, multimodal systems, safety and security, governance
\end{IEEEkeywords}

\input{sections/01_introduction}
\input{sections/02_background}
\input{sections/02.5_arch}
\input{sections/03_architecture}
\input{sections/04_lifecycle}
\input{sections/05_security}
\input{sections/06_discussion}
\input{sections/07_limitations}

\bibliographystyle{IEEEtran}
\bibliography{references/cua_refs}

\end{document}