\section{Introduction}
\label{sec:intro}

Computer-use agents (CUAs) are becoming one of the clearest expressions of agentic AI. Rather than stopping at language generation, these systems inspect interface state, interpret user goals, and execute actions in websites, desktop software, and mobile applications. Recent progress in multimodal modeling, large-scale computer-use data, and trajectory-selection methods has therefore pushed CUAs from narrow automation settings toward more general software operation~\cite{wang2025opencua,gonzalez2025scalingcua}.

This shift changes both capability and risk. In dialogue systems, many errors remain textual. In computer-use settings, the same error may become an unintended purchase, a deleted file, a privilege change, an unsafe command, or a persistent workflow misconfiguration. The challenge becomes sharper as evaluation moves from controlled benchmarks such as Mind2Web and WebArena to open environments such as OSWorld, where agents must cope with unfamiliar interfaces, long-horizon tasks, visual ambiguity, and cross-application dependencies~\cite{deng2023mind2web,zhou2023webarena,xie2024osworld}. Computer use is therefore no longer only a benchmark problem; it is a systems, security, and governance problem.

Open deployment efforts make this transition especially visible. Public descriptions of \textbf{OpenClaw}, for example, suggest a local-first assistant that can be reached through persistent channels such as WhatsApp, Telegram, Slack, Discord, and iMessage~\cite{openclaw2026website,ibm2026openclaw}. Whether OpenClaw becomes a dominant platform is less important than what it reveals analytically: once a CUA is embedded in live channels, tools, memory, sessions, and permissions, capability can no longer be studied independently from authority, trust boundaries, and operational control.

Throughout the paper, we therefore use OpenClaw as a \emph{running example} of open CUA deployment rather than as the sole empirical object of study. The goal is not to claim complete knowledge of its internal implementation, but to use publicly described properties---persistent channel ingress, local-first deployment, tool and session binding, and agent-to-agent interaction---to make the architecture--lifecycle framework concrete in a recognizable open-system setting.

Existing work illuminates important parts of this landscape, but mostly in parallel. GUI-agent and OS-agent surveys clarify architectural design choices, benchmark construction, and platform coverage~\cite{nguyen2025gui,hu2025agents,tang2025survey}. Recent system-building efforts such as OpenCUA and scaling studies such as bBoN show that the field is moving from isolated prototypes toward open foundations and reliability-oriented evaluation~\cite{wang2025opencua,gonzalez2025scalingcua}. Safety benchmarks such as OS-Harm, SafeArena, WAInjectBench, and the more deployment-grounded Risky-Bench make action-level misuse measurable, but they do not by themselves explain how those failures arise from system design or deployment configuration~\cite{kuntz2025harm,tur2025safearena,liu2025wainjectbench,zheng2026riskybench}. Broader agent-security surveys remain useful as governance context rather than primary evidence for CUA-specific architecture and lifecycle claims~\cite{adapala2025aegis,lazer2026survey}. The present survey therefore brings these strands into a single CUA-centered synthesis.

This survey argues that CUAs are best understood through a \emph{joint architecture--lifecycle perspective}. Architecture explains \emph{how} a CUA works: how it reconstructs interface state, forms task-conditioned decisions, and executes actions. Lifecycle explains \emph{when} capabilities, risks, and controls are created, exposed, stressed, and revised. Our central claim is that many practical CUA failures cannot be explained by either lens alone. Capability bottlenecks may originate during training, surface only after deployment, compound during operation, and reappear after updates. A useful survey framework must therefore connect system structure to system evolution over time.

We study general-purpose CUAs across web, desktop, and mobile environments, treating them as end-to-end interactive systems rather than isolated benchmark models. The focus is on how perception, reasoning, action interfaces, execution authority, and operating context interact in realistic deployments.

\paragraph{Survey scope and selection criteria.}
This paper is written as a synthesis survey rather than an exhaustive catalog. We organize the corpus around representative systems, benchmarks, deployment efforts, and security studies available by April 2026, then expand coverage through closely related work cited by those anchor papers. Priority is given to work that either introduces a general-purpose CUA system, benchmark, or deployment setting; clarifies architectural choices in perception, decision-making, or execution; evaluates safety or adversarial misuse in action-taking agents; or addresses deployment, tooling, and governance issues that materially affect real-world CUA operation. Because the area is evolving quickly, the surveyed corpus is weighted toward 2023--2026 work, while earlier papers are retained only when they establish durable concepts such as reflective control, partial observability, or interface-level action modeling. Dialogue-only agents, traditional rule-based robotic process automation, and embodied robotics are discussed only when they directly inform CUA design, evaluation, or governance.

Accordingly, this survey makes three contributions:
\begin{itemize}
  \item \textbf{A joint architecture--lifecycle framework.} We organize the CUA design space using a tri-layer architecture of \emph{Perception}, \emph{Decision}, and \emph{Execution} linked to a four-stage lifecycle of \emph{Creation}, \emph{Deployment}, \emph{Operation}, and \emph{Maintenance}. This framework identifies where capability bottlenecks are formed, where failures first surface, and where controls can be attached.
  \item \textbf{A lifecycle-aware security analysis.} We map the security landscape of CUAs to concrete failure points across the lifecycle, covering prompt injection, interface deception, tool and privilege misuse, memory compromise, agent-to-agent propagation, and supply-chain failures.
  \item \textbf{A governance synthesis for open deployment.} We derive governance implications for realistic computing environments, emphasizing constrained action interfaces, provenance, permission and sandboxing controls, auditing, regression evaluation, and accountability across stakeholders.
\end{itemize}

\begin{table*}[!t]
\centering
\scriptsize
\caption{Representative CUA landscape used to anchor the survey.}
\label{tab:representative_landscape}
\setlength{\tabcolsep}{3.5pt}
\renewcommand{\arraystretch}{1.08}
\begin{tabularx}{\textwidth}{L{1.95cm}L{1.45cm}L{1.55cm}L{2.75cm}L{1.75cm}Y}
\toprule
\textbf{Work} & \textbf{Type} & \textbf{Setting} &
\textbf{Obs./control} & \textbf{Pressure} &
\textbf{Role in this survey} \\
\midrule
Mind2Web~\cite{deng2023mind2web} & Benchmark & Web &
Structured browser state and instruction following & Creation / eval. &
Anchors web grounding and diverse web-task evaluation. \\
WebArena~\cite{zhou2023webarena} & Benchmark & Web &
Reproducible web environments with browser-side control & Operation &
Represents long-horizon web interaction in controlled yet realistic settings. \\
OSWorld~\cite{xie2024osworld} & Benchmark & Desktop / OS &
Open GUI state under real computer tasks & Deploy. / oper. &
Makes visible the shift from browser agents to open computer use. \\
OpenCUA~\cite{wang2025opencua} & Open foundation & Cross-OS &
Demonstration capture, large-scale trajectories, and open end-to-end CUA models & Creation &
Represents the shift from benchmark-only progress toward open data and model foundations for CUAs. \\
UFO~\cite{zhang2025ufo} & System & Desktop &
Hierarchical application routing plus local action execution & Decision / deploy. &
Shows how decomposition reshapes orchestration and multi-app workflows. \\
OS-Copilot~\cite{wu2024copilot} & System & Desktop / OS &
Hybrid GUI interaction and code/tool execution & Exec. / privilege &
Represents the move from pure GUI clicking to higher-authority hybrid action spaces. \\
OS-Harm~\cite{kuntz2025harm} & Safety benchmark & Desktop / OS &
Risk-focused evaluation for action-taking agents & Creation / release &
Captures why action-level safety must be evaluated separately from task success. \\
SafeArena~\cite{tur2025safearena} & Safety benchmark & Web &
Real-website misuse and harmful-task completion & Oper. security &
Represents the misuse-focused evaluation line in realistic web settings. \\
OpenClaw~\cite{openclaw2026website,ibm2026openclaw} & Open deployment exemplar & Multi-channel / local-first &
Persistent ingress, tool binding, sessions, and ecosystem participation & Deploy. / maint. &
Provides the running example for open deployment, governance, and capability diffusion. \\
\bottomrule
\end{tabularx}
\end{table*}

Table~\ref{tab:representative_landscape} previews the representative landscape that anchors the remainder of the survey. Together, these contributions reposition CUAs as lifecycle-dependent interactive systems rather than a standalone benchmark category. The rest of the paper fixes the problem setting and framework, then surveys architectural and lifecycle design choices, maps the security landscape, and distills practical implications for defense and governance.
