\section{Security and Privacy Analysis}
\label{sec:security}

This section analyzes the security and privacy risks associated with computer-use agents (CUAs). Relative to conventional LLM safety, which has largely focused on dialogue behavior and text generation, CUA security must additionally account for \emph{action-level} risk: these systems perceive visual environments, interpret user goals, and execute low-level operations over external interfaces, thereby exposing vulnerabilities at the intersection of perception, reasoning, and execution~\cite{jones2025systematization,foerster2026camels,qian2026zero}.

Following the lifecycle perspective introduced in Section~\ref{sec:lifecycle}, we organize the discussion according to the stages at which security and privacy failures are introduced, amplified, or realized. We first specify a threat model, then introduce an attribution framework that distinguishes \emph{scope overreach}, \emph{objective corruption}, and \emph{environmental misbinding}, and finally review representative threats across creation, deployment, operation, and maintenance. Table~\ref{tab:cua_threats} summarizes this threat classification by lifecycle stage and dominant failure class. Because this paper is motivated in part by open deployment settings illustrated by OpenClaw, we place particular emphasis on risks that emerge when CUA capabilities are exposed through persistent communication channels, extensible tools, and long-lived runtime contexts.

\begin{table*}[!t]
\centering
\scriptsize
\caption{Threat classification by lifecycle stage and dominant failure class (SO: scope overreach; OC: objective corruption; EM: environmental misbinding).}
\label{tab:cua_threats}
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.08}
\begin{tabular}{>{\raggedright\arraybackslash}p{1.7cm}
                >{\raggedright\arraybackslash}p{3.95cm}
                >{\raggedright\arraybackslash}p{1.35cm}
                >{\raggedright\arraybackslash}p{8.0cm}}
\toprule
\textbf{Stage} & \textbf{Threat Type} & \textbf{Attribution} & \textbf{Representative Work} \\
\midrule
Creation & PoT backdoors, training poisoning & OC & BadVLA~\cite{zhou2025badvla}, Hidden Ghost Hand~\cite{cheng2025hidden} \\
Creation & Unsafe action interfaces & EM & OS-Harm~\cite{kuntz2025harm}, CUAHarm~\cite{tian2025measuring} \\
Deployment & Over-privilege, weak isolation & EM & CaMeLs~\cite{foerster2026camels}, CSAgent~\cite{gong2025secure} \\
Deployment & Supply-chain risk (skills, MCP) & OC & Les Dissonances~\cite{li2025dissonances}, MCP Security Bench~\cite{zhang2025mcp} \\
Operation & TOCTOU, action rebinding & EM & Zero-Permission Manipulation~\cite{qian2026zero}, AgentHazard~\cite{liu2025hijacking} \\
Operation & Clickjacking, visual anchors & EM & Chameleon~\cite{zhang2026realistic}, WAInjectBench~\cite{liu2025wainjectbench} \\
Operation & IPI, EIA & OC & InjecAgent~\cite{zhan2024injecagent}, WASP~\cite{evtimov2025wasp}, Chameleon~\cite{zhang2026realistic} \\
Operation & Memory / RAG poisoning & OC & AgentPoison~\cite{chen2024agentpoison} \\
Operation & Scope overreach, PII leakage & SO & AgentDAM~\cite{zharmagambetov2025agentdam}, SafeArena~\cite{tur2025safearena} \\
Operation & Agent-to-agent propagation & OC & Chen~\cite{chen2026openclaw}, Prompt Infection~\cite{lee2024prompt} \\
Maintenance & Governance gaps, stale controls & EM & Risky-Bench~\cite{zheng2026riskybench}, ClawKeeper~\cite{liu2026clawkeeper} \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Threat Model}
\label{subsec:threat_model}

To structure the discussion, we adopt a threat-modeling perspective that identifies attacker roles, protected assets, trust boundaries, platform assumptions, and the major classes of threats emphasized in the recent literature.

\textbf{Attackers.} Prior work suggests at least three recurring attacker roles. First, \emph{external attackers} may control web pages, documents, search results, screenshots, or Model Context Protocol (MCP) tool returns later consumed by the agent. Second, \emph{supply-chain attackers} may distribute malicious or vulnerable skills, plugins, wrappers, or tool integrations. Third, \emph{malicious capability providers} may intentionally publish trojanized services or skills to agent ecosystems. These categories are not mutually exclusive, and some attacks combine multiple layers of compromise.

\textbf{Assets.} The primary assets include: (i) the \emph{confidentiality} of user data such as emails, credentials, files, and personally identifiable information; (ii) the \emph{integrity} of task execution, i.e., whether the agent’s actions remain faithful to user intent; and (iii) control over \emph{privilege-bearing interfaces}, including operating-system capabilities, online accounts, and external services. Compromise along any of these dimensions may result in privacy loss, unauthorized action, or persistent exposure.

\textbf{Trust boundaries.} A core challenge for CUAs is that user intent, environmental content, memory state, and tool-mediated outputs are often processed within a shared reasoning loop. The first trust boundary separates \emph{user-authored intent} from \emph{runtime content}, including interface text, web content, screenshots, retrieved snippets, and documents that may contain hidden or misleading instructions. The second separates \emph{trusted interfaces and tools} from \emph{untrusted external sources}, such as third-party web pages, documents, or MCP servers whose outputs may be incorrect, adversarial, or poisoned. When these boundaries are weakly represented or insufficiently enforced, spoofing, tampering, and privilege escalation become easier to realize.

\textbf{Platform scope.} Risk also varies across \emph{single-agent} and \emph{multi-agent} platforms. In multi-agent settings, prompts, capabilities, and learned behaviors may propagate across agents with limited human review, increasing the difficulty of attribution, containment, and audit. This consideration is especially relevant in open deployment settings, including OpenClaw-like ecosystems, where persistent communication channels, extensible tools, and agent-to-agent interaction can blur the distinction between local execution, external content, and socially propagated capability.

\textbf{Assumptions.} We assume that the underlying model can be induced, misdirected, or manipulated by external content; that the execution environment (web, desktop, or mobile) is only partially trusted; and that the available action space includes high-impact operations such as payment, deletion, installation, authentication, or command execution. Under these assumptions, relatively small failures in intent grounding or interface isolation may produce disproportionate downstream consequences.

\textbf{Threat types.} Following prior systems work, CUA-specific risks can be related to STRIDE-style categories. \emph{Spoofing} includes Indirect Prompt Injection (IPI) and Environmental Injection Attack (EIA), where attacker-controlled content is interpreted as instruction. \emph{Tampering} includes memory poisoning, Retrieval-Augmented Generation (RAG) poisoning, and training-time backdoors. \emph{Repudiation} concerns incomplete provenance, weak audit trails, and insufficient action justifications. \emph{Information disclosure} includes leakage of credentials, documents, and Personally Identifiable Information (PII) across tools or applications. \emph{Denial of service} includes hijacking and resource-exhaustion behaviors that prevent correct task completion. \emph{Elevation of privilege} includes over-privileged interfaces, weak isolation, and action rebinding.

From a lifecycle perspective, these threats manifest differently across stages. During \emph{Creation}, adversaries primarily target training data, model internals, and action abstractions, yielding backdoors or unsafe interfaces. During \emph{Deployment}, missing provenance, weak isolation, or excessive permissions enlarge the practical attack surface. During \emph{Operation}, open and adversarial environments create opportunities for interface deception, prompt injection, and memory poisoning that directly affect confidentiality, integrity, and availability. During \emph{Maintenance}, continuous updates, capability sharing, and ecosystem expansion introduce propagation, regression, and governance challenges. The attribution framework below refines this view by classifying the \emph{dominant operational failure mode} rather than speculating about philosophical intent.

\subsection{Attribution Framework: Dominant Failure Mode}
\label{subsec:attribution}

A recurring issue in the agent-security literature is that unsafe behavior cannot always be attributed in the same way. To clarify this distinction, we adopt a three-part framework based on the \emph{dominant failure mode} that makes an unsafe action operational.

\textbf{Scope overreach} refers to cases in which the agent departs from the user’s intended task boundary without clear evidence that an attacker-supplied objective has become the operative goal. Examples include exploring irrelevant interfaces, collecting unnecessary data, or following salient but non-essential content. The operational signature is off-task behavior under weak task-boundary discipline.

\textbf{Objective corruption} refers to cases in which an attacker-supplied objective becomes operational inside the agent’s action loop. This may happen through injection, poisoning, malicious tools, or backdoor triggers. The operational signature is that the agent begins optimizing for exfiltration, sabotage, or unauthorized redirection rather than merely drifting off-task.

\textbf{Environmental misbinding} refers to cases in which the intended action is remapped by the environment or by weak system binding even though the agent’s immediate objective has not obviously changed. Representative cases include time-of-check to time-of-use (TOCTOU) mismatches, clickjacking, weak perception--action binding, and over-broad authority surfaces. The operational signature is misexecution caused by interface or mediation conditions.

This tripartite framing is intended as an engineering convenience rather than a claim about philosophical agency. It is most useful when applied to the \emph{dominant} failure class that best explains how an unsafe outcome became possible, while still allowing that some incidents span more than one class. In the following subsections, each threat class is discussed in lifecycle context and annotated as scope overreach, objective corruption, or environmental misbinding where appropriate.

\paragraph{Operational decision rule.}
In practice, we apply the framework in three steps. First, ask whether attacker-controlled content, memory state, tooling, or training signal has become the agent's operative objective; if so, the dominant class is \textbf{objective corruption}. Second, if the user-facing goal remains intact but the observation--action binding, privilege mediation, or execution environment redirects the outcome, the dominant class is \textbf{environmental misbinding}. Third, if neither of those conditions holds and the main problem is that the agent has drifted beyond the user's task boundary or collected unnecessary information, the dominant class is \textbf{scope overreach}. This rule does not eliminate all ambiguity, but it makes the attribution scheme sufficiently concrete for lifecycle comparison and control design.

\subsection{Threats in the Creation Stage}
\label{subsec:threats_creation}

Creation-stage threats originate in model design, training data, alignment procedures, or action-interface construction. Although such vulnerabilities may remain latent until deployment or operation, their root causes lie upstream in how the CUA is built. Existing studies mainly emphasize two themes: adversarial manipulation of model internals and unsafe exposure of action primitives.

\subsubsection{Plan-of-Thought Backdoors and Training-time Poisoning}

Within the creation stage, training-time poisoning and Plan-of-Thought (PoT) backdoors are most naturally attributed to \textbf{objective corruption}. In these attacks, adversaries implant triggers into vision-language-action models such that apparently benign inputs induce harmful actions at inference time while nominal performance on clean inputs remains largely intact. BadVLA~\cite{zhou2025badvla}, for example, shows that poisoned training can redirect agent behavior toward attacker-specified endpoints with near-perfect success rates. Hidden Ghost Hand~\cite{cheng2025hidden} further demonstrates composite triggers for mobile GUI agents, reporting attack accuracies as high as 99.7\%. Agent Security Bench~\cite{zhang2024agent} formalizes PoT backdoors as a distinct agent-specific threat class. Taken together, this line of work suggests that harmful objectives can be embedded during model construction yet remain dormant until triggered later in execution.

\subsubsection{Unsafe Action Interfaces and Design Gaps}

A second creation-stage theme concerns \textbf{environmental misbinding} risk arising from unsafe action abstractions. Prior systems analyses argue that low-level, high-entropy action spaces are inherently harder to secure, whereas better interface abstractions can reduce the range of dangerous behaviors exposed to model output~\cite{jones2025systematization}. When high-impact operations such as payment confirmation, deletion, installation, or command execution are delegated too directly, the resulting system becomes more susceptible to spoofing, misexecution, and privilege escalation. Benchmarks such as OS-Harm~\cite{kuntz2025harm} and CUAHarm~\cite{tian2025measuring} further indicate that safety measurements are highly sensitive to interface design choices. From this analytical perspective, these studies point to interface construction itself as part of the security problem, rather than merely a downstream implementation detail.

\subsection{Threats in the Deployment Stage}
\label{subsec:threats_deployment}

Deployment-stage threats emerge when CUAs are connected to real interfaces, tools, services, and user environments. In many cases, these threats do not create entirely new failure modes by themselves; instead, they amplify or enable downstream operation-stage attacks by weakening provenance, isolation, or permission boundaries.

\subsubsection{Over-privilege and Weak Isolation}

The literature generally treats over-privilege and weak isolation as \textbf{environmental misbinding} failures introduced during deployment. Foerster et al.~\cite{foerster2026camels} and Jones et al.~\cite{jones2025systematization} identify missing provenance, weak interaction binding, and unclear privilege boundaries as recurring system-level risks for computer-use agents. When provenance is missing, the agent cannot reliably distinguish user intent from attacker-controlled instruction; when perception and execution are weakly bound, clickjacking and action rebinding become easier; when privilege boundaries are ambiguous, the agent can overreach into sensitive applications or accounts. These concerns become sharper in OpenClaw-like deployments, where persistent sessions, multi-channel ingress, and extensible tool bindings concentrate heterogeneous authority behind a single endpoint. Your Agent, Their Asset~\cite{wang2026asset} makes this pressure concrete for OpenClaw by showing that broad local privileges and persistent state create a large attack surface once live services and files are in scope. Taken together, these studies point to least privilege, sandboxing, and explicit trust separation as baseline deployment mechanisms rather than optional hardening features.

\subsubsection{Supply-chain Risk in Skills and MCP Tools}

Supply-chain risk in skills and MCP-integrated tools is treated here as dominantly \textbf{objective corruption} because the most security-relevant cases involve externally supplied capabilities that replace or contaminate the task the agent should pursue. MCP Security Bench~\cite{zhang2025mcp} systematizes twelve attack types across tool discovery, argument passing, and result return, while Les Dissonances~\cite{li2025dissonances} documents real-world tools capable of confidential-data harvesting, hijacking, and cross-tool pollution. Environmental misbinding remains an important secondary mechanism when tools are not intentionally malicious but still remain exploitable because of excessive permissions, inadequate isolation, or unsafe assumptions about tool outputs. The main lesson is that deployment-time integrity cannot be reduced to malware screening alone: secure composition also depends on capability scoping, provenance tracking, and treating tool outputs as potentially adversarial inputs. OpenClaw-like deployments make this issue especially visible because persistent channels, tool bindings, and longer-lived sessions collapse multiple trust boundaries into one operational surface. ClawKeeper~\cite{liu2026clawkeeper} reaches the same conclusion from the defensive side by treating skills and plugins as first-class security layers rather than incidental extensions.

\subsubsection{Pre-release Red Teaming}

Another deployment-stage concern is weak adversarial evaluation before release. This is best viewed as an \textbf{environmental misbinding} gap in the deployment pipeline. RedTeamCUA~\cite{liao2025redteamcua} shows that hybrid web--OS environments expose interaction paths simplified benchmarks miss. Without such testing, systems may enter deployment with vulnerabilities discovered only after exposure. From a lifecycle viewpoint, pre-release red teaming links creation and maintenance by revealing design flaws early and seeding later regression suites.

\subsection{Threats in the Operation Stage}
\label{subsec:threats_operation}

The operation stage is where most CUA threats become operational. At this stage, models interact with open environments, user data, dynamic interfaces, and external tools under real-time constraints. Accordingly, the operation-stage literature spans the full attribution spectrum, including environmental misbinding, scope overreach, and objective corruption.

\subsubsection{TOCTOU and Action Rebinding}

TOCTOU and action-rebinding attacks are typically classified as \textbf{environmental misbinding}. The central issue is that the agent’s perceptual state and its subsequent action may become desynchronized. Qian et al.~\cite{qian2026zero} report a ``reasoning gap'' of 4--15 seconds between screen understanding and action execution, during which an attacker can alter interface state and redirect the intended click toward a malicious target. Their study demonstrates 100\% authorization hijacking across six Android agents. Similarly, AgentHazard~\cite{liu2025hijacking} shows that deceptive overlays and misleading content can steer agents at non-trivial success rates. In this class of attack, the agent’s local objective may still be correct, while the unsafe outcome is produced by weakened binding between observation and actuation.

\subsubsection{Clickjacking and Visual Anchor Injection}

A related line of work studies \textbf{environmental misbinding} attacks on visual perception. Rather than manipulating reasoning directly, these attacks perturb the perceptual cues on which action selection depends. Chameleon~\cite{zhang2026realistic} shows that optimized visual triggers occupying only a small portion of the screen can substantially increase attack success rates. WAInjectBench~\cite{liu2025wainjectbench} further indicates that textual filtering is insufficient against imperceptible or cross-modal perturbations. Collectively, these studies suggest that CUA security cannot be understood solely as a prompt-level or text-level alignment problem; visual anchoring, layout semantics, and perception robustness are also part of the attack surface.

\subsubsection{Indirect Prompt Injection and Environmental Injection}

Indirect Prompt Injection (IPI) and Environmental Injection Attacks (EIA) occupy a central place in the operation-stage literature. Under the decision rule above, we treat them as dominantly \textbf{objective corruption} because the key transition is that attacker-supplied instructions become operational inside the agent loop. The common mechanism is that instructions embedded in web pages, retrieved documents, search results, screenshots, or MCP tool returns are processed as though they were authoritative. InjecAgent~\cite{zhan2024injecagent}, WASP~\cite{evtimov2025wasp}, SafeSearch~\cite{dong2025safesearch}, Chameleon~\cite{zhang2026realistic}, and GhostEI-Bench~\cite{chen2025ghostei} together show that this threat spans web, dynamic, and mobile environments. PASB~\cite{wang2026pasb} extends the point to personalized OpenClaw deployments, where prompt processing, tool usage, and memory retrieval remain jointly vulnerable over long horizons. Across these studies, environmental content and tool outputs should be treated as potentially adversarial inputs rather than neutral context~\cite{zhang2025browsesafe}.

\subsubsection{Memory and RAG Poisoning}

Memory poisoning and RAG poisoning are generally attributed to \textbf{objective corruption}. Here the attacker seeks not merely to redirect a single episode, but to implant persistent harmful influence into the retrieval or memory substrate on which future behavior depends. AgentPoison~\cite{chen2024agentpoison} and Agent Security Bench~\cite{zhang2024agent} already show that memory compromise can be more persistent than one-shot injection. Les Dissonances~\cite{li2025dissonances} complements this perspective through cross-tool harvesting and pollution, while Your Agent, Their Asset~\cite{wang2026asset} sharpens the concern for personalized OpenClaw deployments by showing that poisoning persistent capability, identity, or knowledge state can substantially raise attack success rates. The broader lesson is that the unit of compromise shifts from the immediate prompt to the longer-lived informational state that shapes future reasoning.

\subsubsection{Scope Overreach and Data Minimization Violations}

Not all privacy failures arise from explicit adversarial compromise. A growing body of work instead highlights \textbf{scope overreach}, in which agents exceed task scope, inspect irrelevant interfaces, or collect unnecessary data. AgentDAM~\cite{zharmagambetov2025agentdam} shows that mainstream agents often fail to satisfy data-minimization expectations and may capture or leak irrelevant PII. By contrast, \textbf{objective corruption} is the better description when the agent has been induced to exfiltrate or misuse data under attacker influence. SafeArena~\cite{tur2025safearena} finds that GPT-4o can complete a substantial fraction of harmful tasks on real websites, and prior systems work suggests that tool chaining may facilitate data movement across applications~\cite{jones2025systematization,li2025dissonances}. The literature therefore points to a dual requirement: task-boundary discipline is needed to limit benign overcollection, while adversarial robustness is needed to prevent attacker-directed exfiltration. For the classification used in Table~\ref{tab:cua_threats}, the dominant class is \textbf{scope overreach} whenever off-task collection is the primary mechanism.

\subsubsection{Long-horizon Constraint Forgetting}

Long-horizon operation introduces additional failure modes that are partly \textbf{scope overreach} and partly \textbf{environmental misbinding}. LPS-BENCH~\cite{chen2026lps} suggests that agents often lack sufficient foresight to reason about cascading consequences across multi-step plans, allowing goal completion pressure to dominate safety constraints. Related results from OS-Harm~\cite{kuntz2025harm} and CUAHarm~\cite{tian2025measuring} indicate that dialogue-level alignment does not automatically transfer to action-oriented settings. In aggregate, this literature points to a temporal dimension of CUA risk: safety constraints must remain stable not only within single decisions, but across extended interaction horizons in which partial progress can gradually erode earlier safeguards.

\subsubsection{Agent-to-agent Propagation and Cross-agent Threats}

Cross-agent interaction introduces a distinct operation-stage threat surface. For threat classification purposes, we treat the dominant risk as \textbf{objective corruption}: once harmful prompts, skills, or demonstrations are propagated across agents, foreign objectives can replicate at machine scale. Chen et al.~\cite{chen2026openclaw} provide the first large-scale empirical study of peer learning among AI agents on Moltbook, reporting participation by 2.4M agents, an 11.4:1 ratio of teaching to help-seeking statements, and machine-scale propagation of skills with minimal human review. These findings are especially relevant to open ecosystems illustrated by the publicly discussed OpenClaw/Moltbook example, where persistent communication channels and socially mediated capability sharing can transform local mistakes or poisoned content into ecosystem-level exposure.

Prompt Infection~\cite{lee2024prompt} shows that harmful prompts can replicate across agents in multi-agent systems, while Multi-Agent Security Tax~\cite{peigne2025multi} highlights the trade-off between stronger security controls and collaborative flexibility. Broader agent-security surveys such as Aegis and the Agentic AI Cybersecurity Survey are useful here as context for collusion, cascading failure, and oversight-evasion concerns, but the concrete CUA lesson is narrower: open agent societies require identity verification, provenance-preserving exchange, semantic malware containment, and governance policies for capability diffusion~\cite{adapala2025aegis,lazer2026survey}.

The OpenClaw/Moltbook example is useful here for the same reason. As publicly discussed, it makes explicit that capability propagation may occur through social or ecosystem channels rather than only through direct model updates, which means that provenance and governance must apply to inter-agent exchange as well as to standalone model execution.

\subsection{Threats in the Maintenance Stage}
\label{subsec:threats_maintenance}

Maintenance-stage threats concern long-term operation after release. They are primarily \textbf{environmental misbinding} in character, but remain central to whether CUA deployments stay safe under model updates, interface drift, evolving tools, and expanding ecosystems.

\subsubsection{Governance Gaps and Continuous Evaluation}

A consistent theme in recent work is that CUA security cannot be treated as a one-time property established before deployment. Model updates, tool changes, interface redesign, and policy adjustments may all introduce regressions. Risky-Bench~\cite{zheng2026riskybench} shows that deployment-grounded evaluation reveals risks that narrower task suites can miss. PASB~\cite{wang2026pasb} adds that personalized long-horizon use exposes vulnerabilities across prompt processing, tool usage, and memory retrieval that are easy to understate in synthetic settings. In OpenClaw specifically, Your Agent, Their Asset~\cite{wang2026asset} shows that poisoning persistent capability, identity, or knowledge state can sharply raise attack success rates, while ClawKeeper~\cite{liu2026clawkeeper} argues that fragmented point defenses are insufficient without cross-stage protection over skills, plugins, and watcher-style intervention. The maintenance problem is therefore not merely operational hygiene; it is a structural requirement for CUA ecosystems in which capabilities, interfaces, and trust relationships continue to evolve after release. This observation directly motivates the lifecycle-aligned defenses and governance recommendations discussed in Section~\ref{sec:discussion}.
