\section{Security Principles and Existing Practices}
\label{sec:discussion}

\subsection{Security Implications and Design Principles}
\label{subsec:implications}

The security landscape described in Section~\ref{sec:security} carries broad implications for stakeholders. Building on the lifecycle-oriented threat analysis and attribution framework, this section distills key security principles and existing practices for different roles. We organize them by lifecycle role and attribution focus, aligning with the threat model and the curiosity--malice attribution framework.

\subsubsection{CUA Developers (Creation stage)}
For CUA developers, security must be addressed at design time. \textbf{Threat model and safe interfaces:} A core principle is to \emph{design for least privilege and safe interfaces by default}, adopting the threat model in Section~\ref{subsec:threat_model} and accounting for UI deception~\cite{jones2025systematization}, indirect prompt injection~\cite{zhan2024injecagent,evtimov2025wasp}, and action rebinding~\cite{qian2026zero}. Because threats span curiosity and malice, developers cannot rely on dialogue-level alignment alone. Scenario-specific alignment and behavioral measurement are required~\cite{kuntz2025harm,tian2025measuring}. \textbf{Supply chain:} Hygiene for skills, plugins, and tool definitions is essential. Metadata integrity verification and digital signing should precede distribution. Evil Geniuses~\cite{tian2023evil} and the Microsoft CUA systematization~\cite{jones2025systematization} underscore that agent security combines NLP safety with classic software security.

\subsubsection{Deployment and Integration Teams (Deployment stage)}
For deployment and integration teams, system-level boundaries are paramount. \textbf{Isolation and boundaries:} A guiding principle is to \emph{treat the environment as potentially adversarial} and enforce strong isolation. Over-privilege and weak isolation enable operation-stage attacks. Least-privilege access, sandboxing~\cite{meng2025cellmate,foerster2026camels}, and provenance-aware interfaces must be enforced before agents enter production. \textbf{Gating and red teaming:} The MCP Security Bench~\cite{zhang2025mcp} and Agent Security Bench~\cite{zhang2024agent} show that stronger models are often more exploitable. Deployment gating and pre-release red teaming~\cite{liao2025redteamcua} are therefore essential.

\subsubsection{Operators (Operation stage)}
For operators, the binary attribution (curiosity vs.\ malice) matters. Curiosity-driven overreach requires task-boundary and data-minimization controls. Malice-driven threats (e.g., Indirect Prompt Injection (IPI), memory poisoning) require intent verification and input sanitization. A practical principle is to \emph{close the loop with continuous monitoring and human confirmation for high-impact actions}. Runtime monitoring, misalignment detection~\cite{ning2026actions}, and human-in-the-loop confirmation remain critical. Operators must treat tool outputs as potentially adversarial and enforce constraints on cross-tool composition~\cite{zhang2025browsesafe}.

\subsubsection{Maintainers and Ecosystem Stewards (Maintenance stage)}
For maintainers and ecosystem stewards, agent-to-agent propagation introduces governance gaps. A key principle is to \emph{govern agent societies as first-class security subjects}, not just as implementation details. Platforms such as OpenClaw--Moltbook~\cite{chen2026openclaw} demonstrate that skills propagate at machine scale with minimal vetting. Identity and access management for non-human entities, content governance for agent social platforms, and defenses against Prompt Infection~\cite{lee2024prompt} are emerging priorities. Centralized or federated registries with verified identities, version pinning, and audit trails help mitigate supply-chain risks.

\subsubsection{Researchers}
For researchers, a key finding is that \textbf{safety alignment in dialogue does not automatically transfer to action}. A research principle is to \emph{measure what matters at the action level}: benchmarks such as OS-Harm~\cite{kuntz2025harm}, RiOSWorld~\cite{yang2025riosworld}, and CUAHarm~\cite{tian2025measuring} demonstrate that scenario-specific alignment, environment-aware policies, and behavioral measurement are necessary. Research should prioritize cross-modal defenses (text and vision), plan-time safety awareness~\cite{chen2026lps}, and formal verification~\cite{miculicich2025veriguard}. The attribution framework also highlights an underexplored challenge: \emph{observing and attributing} curiosity vs.\ malice in agent behavior---critical for designing targeted defenses.

\subsection{Recommendations and Defense Frameworks}
\label{subsec:recommendations}

Defense research is shifting from input filtering to lifecycle-spanning, multi-layer protection. We summarize representative frameworks by lifecycle stage and attribution target.

\subsubsection{Creation stage}
\emph{Formal verification} (VeriGuard~\cite{miculicich2025veriguard}) applies provably sound verification to code-agent outputs, achieving 0\% ASR on evaluated benchmarks. \emph{Safe interfaces and abstractions} reduce low-level action entropy. Constrained primitives force high-impact actions into auditable operations. \emph{Benchmarks} (OS-Harm, CUAHarm, LPS-BENCH) convert safety from qualitative claims into quantitative signals for release gating.

\subsubsection{Deployment stage}
\emph{Intent-aware access control} (CSAgent~\cite{gong2025secure}) binds actions to task intent via contextual integrity, defending against 99.36\% of attacks with ~7\% latency overhead. \emph{Sandboxing} (ceLLMate~\cite{meng2025cellmate}, CaMeLs~\cite{foerster2026camels}) constrains the blast radius of prompt injection via isolation and blind planning. \emph{Pre-release red teaming} (RedTeamCUA~\cite{liao2025redteamcua}) stress-tests agents before production.

\subsubsection{Operation stage}
\emph{Intent provenance and misalignment detection} (IntentGuard~\cite{kang2025mitigating}, DeAction~\cite{ning2026actions}) trace action intent to its source and detect off-task or misaligned actions, cutting attack success by over 90\% without harming benign performance. \emph{Reasoning-layer guardrails} (MirrorGuard~\cite{zhang2026mirrorguard}) correct unsafe reasoning in simulation before GUI execution, reducing unsafe rates from 66.5\% to 13\%. \emph{Privacy protection} (PrivacyChecker~\cite{wang2025privacy}) filters agent-to-agent communication using context integrity, reducing PII leakage from 36\% to 7--8\%. Data minimization should be evaluated behaviorally~\cite{zharmagambetov2025agentdam}. \emph{Advanced guardrails} (ShieldAgent~\cite{chen2025shieldagent}, AgentDoG~\cite{liu2026agentdog}) provide real-time monitoring and diagnostic classification of risk sources.

\subsubsection{Maintenance stage}
\emph{Governance for agent societies} (Aegis Protocol~\cite{adapala2025aegis}) proposes decentralized identifiers, post-quantum cryptography, and zero-knowledge proofs for verifiable policy compliance. LLM Tagging~\cite{lee2024prompt} mitigates infection spread in multi-agent systems. Continuous evaluation, regression suites, and vulnerability management operationalize benchmarks into release gates and incident response.

No single defense suffices. A layered approach combining intent verification, reasoning correction, formal methods, sandboxing, and governance is needed. Developers and operators should select defenses based on threat model, lifecycle stage, attribution class (curiosity vs.\ malice vs.\ environmental), and acceptable performance-cost trade-offs.
